---
title: "linear model from Web"
author: "Sierra Lander"
date: "Summer 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS


##Methods 

all variables are numerical so we use num~num   

### Descriptive results 

### scatterplot

```{r}
plot(SIMS~ARM,data=data)
```

The plot shows that the arm strength goes up and so does sims.   

### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```



#### scatterplot with model fit


```{r}
plot(SIMS~ARM,data=data)
abline(model.1, col='red')
```


The p-values for the intercept and slope are  0.00000000000000022   


### Numerical results



```{r}
cor(SIMS~ARM,data=data)
```
the correlation shows a positive number which means there is a positive correlation



#### Predict at target point

```{r}
newdata = data.frame(ARM = 88, GRIP = 94)
predict(model.1, newdata, interval='pred')
```


The SIMS prediction for ARM=88 is 0.71, the prediction interval is -1.73, 3.14)  
  
  
  
  
  
  
  
  
  
  

## Model 2 SIM~GRIP

##Methods 

All variables are numerical so we use file num~num



### Now add in scatterplot

```{r}
plot(SIMS~GRIP,data=data)
```


The graph shows that as grip goes up so does sims 



#### predict model.2 at target point

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```



### Inferential  (Build model.2)

```{r}
plot(SIMS~GRIP,data=data)
abline(model.2, col='red')
```


The p-values for the intercept and slope are 0.00000000000000022  




### Numerical results 

```{r}
cor(GRIP~SIMS,data=data)
```

The correlation is positive because it shows a positive number



## Predict at targer point


```{r}
newdata = data.frame(SIMS = 95, GRIP = 94)
predict(model.2, newdata, interval='pred')
```


The SIMS prediction for GRIP=94 is 0.54, the prediction interval is -3.10,2.04)  
  

## Model 3 SIM~ARM+GRIP

prediction interval of SIMS Given a value of 94 for GRIP and 88 for ARM

both variables are numerical so we use helper file num~num


### Numerical results (cor)

```{r}
df = data.frame(SIMS=data$SIMS, GRIP=data$GRIP, ARM=data$ARM)
cor(df)
```

correlation matrix shows that SIMS correlates with GRIP and ARM, and GRIP correlates with SIMS
  
### Inferential  (Build 2-dimentional model.3)
```{r}
model.3 = lm(SIMS~ ARM + GRIP, data=data)
summary.lm(model.3)
```

p-value from the F-statistic is  0.00000000000000022 this means that the data set fits the model


p-vales for the  parameters: intercept, coefficient of ARM, and Coefficient of Grip are  0.00000000000000022, .00000000150, .00000499

#### predict model.3 at target point
  prediction interval of SIMS given value of 94 for GRIP and 88 for ARM


```{r}
newdata= data.frame(GRIP=94, ARM=88)
predSIMS = predict(model.3, newdata, interval = 'pred')
print(predSIMS)
```


The predicted SIMS score for GRIP=94, ARM=88 is 0.150 

,

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(model.1,model.3)
```

the anvoa shows that there is a difference on the models. the p-value is 0.000004. The RSS in model 2 is 188 which is smaller than model 1 which is 217. Therefore SIMS~ARM+GRIP is the better model because model 1 is SIMS ~ ARM only. 



### Model.2 vs Model.3

```{r}
anova(model.2, model.3)
```

The anova shows that model 2 (1) SIMS~GRIP has a RSS of 243 and Model 3 SIMS~ARM+GRIP(2) has a RSS of 188  this shows that model 3 is the better model of the two and The p-value is .000005. 


## Informally compare Model.1 with model.2
   
```{r}
summary.lm(model.1)
summary.lm(model.2)
```
   
Model 1 sim~arm is better than model 2 sims~grip because is has a smaller residual standard error (1.295, 1.226)
   
   
   
